Towards typed repositories of proofs \\[0.6em] \small \textsf{MIPS 2010}

### 2010-07-09 talks/MIPS2010/slides.tex
bug corrigé ds les illustrations, qqes fixes du handout
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@149 403e655e-980a-47bd-8c60-1ee5348bb6b7

The actual text of the proofs for versioning is often
non-satisfactory: on one side it contains mostly non-relevant
informations (line breaks, spaces\ldots), on the other the relevant
ones, the validity of the proof, the dependency between objects, is
absent.

So both these facilities, among others, seem to provide ad-hoc ways to
deal with the non-linearity of the script and its edition. They
provide a coarse approximation of the dependency amongst objects. But
this notion of dependency \emph{is} a metatheoretical property of a
logic, and it could be used to provide an \emph{exact} way of dealing
with the impact of changes, in particular to guide proof-checking to
recheck only minimal parts of the developments.

@@

This way, we ensure that no two subdirectories with different content
appear in the database, that is, we enforce maximal sharing among
subdirectories. In functional programming, this is called
hash-consing.

@@

\hrule
\begin{frame}{Incremental type-checking, incrementally}%
  \begin{block}{Syntax}%
    \only<1>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ x\ |\ t\ t\ %
      |\ *\] \[\]}%
    \only<2>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ \alert x\ |\ %
      \alert{t\ t}\ |\ *\] \[\]}%
    \only<3>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ \alert a\ |\ %
      *\]%
      \[ a\ ::=\ x\ |\ a\ x \]}%
    \only<4>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ a\ |\ *\ |\ %
      \alert{(x = a)\cdot t}\]%
      \[ a\ ::=\ x\ |\ a\ x \]}%
    \only<5>{\[ t\ ::=\ \alert{[x:t]\cdot t}\ |\ (x:t)\cdot t\ |\ a\ %
      |\ *\ |\ (x = a)\cdot t\]%
      \[ a\ ::=\ x\ |\ a\ x \]}%
    \only<6->{\[ t\ ::=\ (x:t)\cdot t\ |\ a\ |\ *\ |\ (x = a)\cdot t\]%
      \[ a\ ::=\ x\ |\ a\ x \]}%
  \end{block}
  \begin{block}{Environments}
      \[ \Gamma\ ::=\ \cdot\ |\ \Gamma [x:t]\only<4>{\ |\
        \alert{\Gamma[x=a:t]}}
      \only<5->{\ |\ \Gamma[x=a:t]} \] 
    \end{block}
    \begin{block}{Judgement}
      \vspace{1em}
        \begin{tabular}{ll} \Large
          $\Gamma\vdash t : u \uncover<7>{\alert{\Rightarrow\Delta}}$
          &
          \only<-6>{`` In environment $\Gamma$, term $t$ has type $u$
            ''}
          \uncover<7>{`` From
            \alert{repository} $\Gamma$, term $t$ of type $u$ \\ &
            \alert{\hspace{1.2em}leads to the new repository $\Delta$}''
          }
        \end{tabular}
    \end{block}
\end{frame}
\hrule

Let's construct a calculus for this. We start from the usual Pure Type
Systems, with abstraction, products, variables, applications and just
one sort for the sake of simplicity. Typing is done in the usual
name-type environment, by the usual judgement: ``In $\Gamma$, $t$ has
type $u$. We modify it gradually, keeping its expressivity but
tailoring the syntax to name all subterms.

@@

It happens that the system is already interesting enough for our
purpose without abstraction: the notions of dependency and
incrementality do not need the abstraction. For the moment, we remove
it, and with it all notion of computation.

>>>>

So both these facilities provide a coarse approximation of the
dependency between objects. But this notion of dependency \emph{is} a
metatheoretical property of a logic, and it could be used to provide
an \emph{exact} way of dealing with the impact of changes, in
particular to guide proof-checking to recheck only minimal parts of
the developments.

@@

This way, we ensure that no two subdirectories with the same content
appear in the database, that is, we enforce maximal sharing among
subdirectories. In functional programming, this is called
hash-consing.

@@

\hrule
\begin{frame}{Incremental type-checking, incrementally}%
  \begin{block}{Syntax}%
    \begin{overlayarea}{\textwidth}{3.5em}
      \only<1>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ x\ |\ t\
        t\ %
        |\ *\] \[\]}%
      \only<2>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ \alert x\
        |\ %
        \alert{t\ t}\ |\ *\] \[\]}%
      \only<3>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ \alert a\
        |\ %
        *\]%
        \[ a\ ::=\ x\ |\ a\ x \]}%
      \only<4>{\[ t\ ::=\ [x:t]\cdot t\ |\ (x:t)\cdot t\ |\ a\ |\ *\
        |\ %
        \alert{(x = a)\cdot t}\]%
        \[ a\ ::=\ x\ |\ a\ x \]}%
      \only<5>{\[ t\ ::=\ \alert{[x:t]\cdot t}\ |\ (x:t)\cdot t\ |\
        a\ %
        |\ *\ |\ (x = a)\cdot t\]%
        \[ a\ ::=\ x\ |\ a\ x \]}%
      \only<6->{\[ t\ ::=\ (x:t)\cdot t\ |\ a\ |\ *\ |\ (x = a)\cdot
        t\]%
        \[ a\ ::=\ x\ |\ a\ x \]}%
    \end{overlayarea}
  \end{block}
  \begin{block}{Environments}
      \[ \Gamma\ ::=\ \cdot\ |\ \Gamma [x:t]\only<4>{\ |\
        \alert{\Gamma[x=a:t]}}
      \only<5->{\ |\ \Gamma[x=a:t]} \] 
    \end{block}
    \begin{block}{Judgement}
      \vspace{1em}
        \begin{tabular}{ll} \Large
          $\Gamma\vdash t : u \uncover<7>{\alert{\Rightarrow\Delta}}$
          &
          \only<-6>{`` In environment $\Gamma$, term $t$ has type $u$
            ''}
          \uncover<7>{`` From
            \alert{repository} $\Gamma$, term $t$ of type $u$ \\ &
            \alert{\hspace{1.2em}leads to the new repository $\Delta$}''
          }
        \end{tabular}
    \end{block}
\end{frame}
\hrule

Let's construct a calculus for this. We start from the usual Pure Type
Systems, with abstraction, products, variables, applications and just
one sort for the sake of simplicity. Typing is done in the usual
name-type environment, by the usual judgement: ``In $\Gamma$, $t$ has
type $u$''. We modify it gradually, keeping its expressivity but
tailoring the syntax to name all subterms.

@@

To understand how incrementality is handled, we focus for the moment
on a minimal system. It happens that it is already interesting enough
without abstraction. For the moment, we remove it, and with it all
notion of computation.
